### 2023-02-15 paper-thesis/chapters/01-introduction.tex
Minor fixes
Some layout fixes

\section*{Contributions}
This goal of this thesis is to investigate how well Statix is fit for the task of defining a simple dependently-typed language. We want to investigate whether typical features of dependently typed languages can be encoded in Statix. The goal is not only to show that Statix can implement it, but also that the implementation is concise.

@@

The source code for this thesis can be found online: \url{https://github.com/JonathanBrouwer/master-thesis}

>>>>

\section*{Contributions}
This goal of this thesis is to investigate how well Statix is fit for the task of defining a dependently-typed language. We want to investigate whether typical features of dependently typed languages can be encoded in Statix. The goal is not only to show that Statix can implement it, but also that the implementation is concise.

@@

The source code for this thesis can be found online \\ \url{https://github.com/JonathanBrouwer/master-thesis}

### 2023-02-15 paper-thesis/chapters/20-adding-booleans.tex
Add chapter transitions
Some layout fixes

Now that a few features have been implemented, we can discuss how easy the language is to extend. The conclusion so far is that it is doable in a clean way. Specifically, the approach for each of the features above was:

\begin{enumerate}
	\item Define the parsing rules for the new feature
	\item Create a new file, \verb|tp_bools.stx| and import this file in the main \verb|type_check.stx| file.
	\item In the new file, define a case for the \verb|betaReduceHead| rule for the constructors that were added. Also define cases for \verb|expectBetaEq| and \verb|betaReduce| if necessary. It is necessary iff the new constructors are not always eliminated by \verb|betaReduceHead|.
	\item In the new file, define a case for the \verb|typeOfExpr| rule for the constructors that were added.
\end{enumerate}

This allows each feature to be isolated to its own file. If we decide that we don't like a feature after all, we can remove it simply by unimporting the file. This problem is similar to the expression problem\cite{expression_problem}, which Statix solves very well.

In the following chapters, we're going to be extending the language further with these more interesting features using this approach:
\begin{itemize}
	\item Inference (chapter~\ref{ch:inference})
	\item Inductive Datatypes (chapter~\ref{ch:datatypes})
	\item Universes (chapter~\ref{ch:universes})
\end{itemize}

>>>>

Now that a few features have been implemented, we can discuss how easy the language is to extend. From the three examples above and the following chapters we can conclude that extending the language in a clean way is possible. To extend the language, we used the following approach:

\begin{enumerate}
	\item Define the parsing rules for the new feature
	\item Create a new file, \verb|tp_[feature].stx| and import this file in the main \verb|type_check.stx| file.
	\item In the new file, define a case for the \verb|betaReduceHead| rule for the constructors that were added. Also define cases for \verb|expectBetaEq| and \verb|betaReduce| if necessary. This is only necessary iff the new constructors are not always eliminated by \verb|betaReduceHead|.
	\item In the new file, define a case for the \verb|typeOfExpr| rule for the constructors that were added.
\end{enumerate}

This allows each feature to be isolated to its own file. If we decide that we don't like a feature after all, we can remove it simply by unimporting the file. This problem is similar to the expression problem\cite{expression_problem}, where we want to extend datatypes and their associated behaviour, which Statix solves very well.

In the following chapters, we're going to be extending the language further with these more interesting features using the approach defined above:
\begin{itemize}
	\item Inference (chapter~\ref{ch:inference})
	\item Inductive Datatypes (chapter~\ref{ch:datatypes})
	\item Universes (chapter~\ref{ch:universes})
\end{itemize}

### 2023-02-15 paper-thesis/chapters/40-discussing-statix.tex
Some layout fixes

Now that we've implemented a feature-rich dependently typed language in Statix, this chapter discusses the experience, and recommends some features which might improve the experience.

@@

	isStrictlyPositive : (scope * Expr)
	isStrictlyPositive(s, FnType(...)) = ...
	\end{lstlisting}
	The definition of \verb|isStrictlyPositive| can then use all the metavariables of \verb|DataTypeDecl| without having to explicitly pass them.

>>>>

Now that we've implemented a feature-rich dependently typed language in Statix, this chapter discusses the experience, and recommends some changes to Statix and Spoofax which might improve the experience.

@@

	isStrictlyPositive : (scope * Expr)
	isStrictlyPositive(s, FnType(...)) = ...
	\end{lstlisting}
	The definition of \texttt{isStrictlyPositive} can then use all the metavariables of \\ \texttt{DataTypeDecl} without having to explicitly pass them.

### 2023-02-15 paper-thesis/chapters/98-relatedwork.tex
Fix chapter references
Some layout fixes

@@

In chapter \ref{ch:comp-haskell} we compared the implementation in Statix with an implementation in Haskell. The pi-forall language~\cite{pi_forall} is a language with a similar complexity to the language presented in this paper. In principle, the implementations are very similar. For example, the inference rules of pi-forall are similar to the inference rules presented in figure \ref{fig:type-check-rules} from this paper. The primary difference is that they use a bidirectional type system~\cite{bidirectional}, whereas this paper uses Statix' unification.

>>>>

In chapter \ref{ch:comp-haskell} and \ref{ch:comp-lambdapi} we already compared our implementation with two different implementations.

@@

In chapter~\ref{ch:comp-haskell} we compared the implementation in Statix with an implementation in Haskell. The pi-forall language~\cite{pi_forall} is a language with a similar complexity to the language presented in this paper. In principle, the implementations are very similar. For example, the inference rules of pi-forall are similar to the inference rules presented in figure \ref{fig:type-check-rules} from this paper. The primary difference is that they use a bidirectional type system~\cite{bidirectional}, whereas this paper uses Statix' unification.
